### 2014-10-24 thesis/method/classification.tex
Added reference to Rosenblatt (1962)

Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
into account equally when classifying a new example. However, since feature
values often span different ranges and units of measurement, a particularly
large range of values for a feature would contribute much more to the distance
than another feature that spans a lower set of values.
To avoid this, it is important to \textit{normalise} the
values of each numeric feature before using $k$-NN: for each feature $i$, we
calculate the normalised value of the feature for each training example. This
is done using the following relationship:
\begin{equation*}
a_i = \dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}
\end{equation*}
where $a_i$ is the normalised value, $v_i$ is the actual value in the data
set, and max $v_i$ and min $v_i$ are taken over all training examples.

@@

\section{Multi-layer Perceptron (MLP)}
Multi-layer perceptrons (MLPs) are a generalised form of the
\textit{perceptron}, first
studied extensively by Rosenblatt \todo{add citation entry}. The original
motivation of the perceptron 
was to develop a very simple artificial neural network to model the
behaviour of neurons in the brain. The perceptron model for a two-class
problem is a linear combination of the features: that is,
$h(\mathbf{x}) = \sum w_i x_i$ where the $w_i$ are the weights of the model.
There is an input node, called a \textit{neuron}, for each feature, and the
weights are applied to the connections between these input neurons to the
output.
Usually, a constant term of 1 called the \textit{bias} is also added to the
model for an additional trainable input. To classify an item, we substitute it
into $h(\mathbf{x})$; if the value is above a particular threshold, we label it
as class 1, else it belongs in class 2. For convenience, we can let this
threshold be 0. Training the perceptron involves the following steps:
\begin{enumerate}
\item Start with random numbers for the weights $w_i$.
\item For each training example $\mathbf{x_i}$: calculate $h(\mathbf{x_i})$.
If it is correctly classified, do not update the weights.
If it is incorrectly classified, add a multiple of the misclassified
instance to the current weights using
$\mathbf{w} = \mathbf{w} + \lambda \mathbf{x_i}$, where $\lambda$ is a small
constant.
\item Repeat the previous step until all training examples are correctly
classified.
\end{enumerate}
If the data are linearly separable, the perceptron is guaranteed to find a
hyperplane that separates the two classes. However, it is not able to do this
when there is a non-linear decision boundary -- for that we will need the
multi-layer perceptron.

>>>>

Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
into account equally when classifying a new example. However, since feature
values often span different ranges and units of measurement, a particularly
large range of values for a feature would contribute much more to the distance
than another feature that spans a lower set of values.
To avoid this, it is important to \textit{normalise} the
values of each numeric feature before using $k$-NN: for each feature $i$, we
calculate the normalised value of the feature for each training example. This
is done using the following relationship:
\begin{equation*}
a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
\end{equation*}
where $a_i$ is the normalised value, $v_i$ is the actual value in the data
set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is a
scaling factor $t$ indicates how much to translate the range by. If $s=1$
and $t=0$, then the range of normalised values will be in $[0,1]$.

@@

\section{Multi-layer Perceptron (MLP)}
Multi-layer perceptrons (MLPs) are a generalised form of the
\textit{perceptron}, first
studied extensively by Rosenblatt \cite{Rosenblatt1962}. The original
motivation of the perceptron 
was to develop a very simple artificial neural network to model the
behaviour of neurons in the brain. The perceptron model for a two-class
problem is a linear combination of the features: that is,
$h(\mathbf{x}) = \sum w_i x_i$ where the $w_i$ are the weights of the model.
There is an input node, called a \textit{neuron}, for each feature, and the
weights are applied to the connections between these input neurons to the
output.
Usually, a constant term of 1 called the \textit{bias} is also added to the
model for an additional trainable input. To classify an item, we substitute it
into $h(\mathbf{x})$; if the value is above a particular threshold, we label it
as class 1, else it belongs in class 2. For convenience, we can let this
threshold be 0. Training the perceptron involves the following steps:
\begin{enumerate}
\item Start with random numbers for the weights $w_i$.
\item For each training example $\mathbf{x_i}$: calculate $h(\mathbf{x_i})$.
If it is correctly classified, do not update the weights.
If it is incorrectly classified, add a multiple of the misclassified
instance to the current weights using
$\mathbf{w} = \mathbf{w} + \lambda \mathbf{x_i}$, where $\lambda$ is a small
constant.
\item Repeat the previous step until all training examples are correctly
classified.
\end{enumerate}
If the data are linearly separable, the perceptron is guaranteed to find a
hyperplane that separates the two classes. However, it is not able to do this
when there is a non-linear decision boundary -- for that we will need the
multi-layer perceptron.
